---
title: "hw2"
output: html_document
author: Lumi Huang
---
Q1
Read Chapter 7 (Exploratory Data Analysis) of R for Data Science and do exercises 7.3.4, 7.4.1, 7.5.1.1, 7.5.2.1, and 7.5.3.1

7.3.4
1.Explore the distribution of each of the x, y, and z variables in diamonds. What do you learn? Think about a diamond and how you might decide which dimension is the length, width, and depth.

```{r}
library(tidyverse)
#examining the distribution of x, y, z
ggplot(data = diamonds) + geom_histogram(mapping = aes(x), binwidth = 0.5)
ggplot(data = diamonds) + geom_histogram(mapping = aes(y), binwidth = 0.5) + 
  coord_cartesian(xlim = c(0, 15))
ggplot(data = diamonds) + geom_histogram(mapping = aes(z), binwidth = 0.5) +
  coord_cartesian(xlim = c(0,10))

#I learnt that there are some outliers in y and z by the evidence of the unusually wide limits on the x-axis. Based on the distribution of x, y, z without outliers, we can say x, y are the length and width of diamonds, and z is the depth of diamonds.

```

2.Explore the distribution of price. Do you discover anything unusual or surprising? (Hint: Carefully think about the binwidth and make sure you try a wide range of values.)
```{r}
#small value for binwidth
ggplot(data = diamonds) + 
  geom_histogram(mapping = aes(x = price), binwidth = 10)

#try a wide range of values
ggplot(data = diamonds) + 
  geom_histogram(mapping = aes(x = price), binwidth = 150)

#The distribution of price is very positively skewed. It looks like there is a lower bound for the price of diamonds based on the distribution of price.
```

3.How many diamonds are 0.99 carat? How many are 1 carat? What do you think is the cause of the difference?
```{r}
carat_0.99 <- diamonds %>%
  filter(carat == 0.99)
count(carat_0.99)

carat_1 <- diamonds %>%
  filter(carat == 1)
count(carat_1)

mean(carat_0.99$price)
mean(carat_1$price)

#there are 23 diamonds that are 0.99 carat, and 1558 diamonds that are 1 carat. The price differences between 0.99 carat and 1 carat diamonds could be the cause of this difference. Because 1 carat diamond is about $800 more expensive than 0.99 carat diamond, we expect to see more 1 carat diamonds than 0.99 carat diamonds even if the other figures are about the same.
```


4.Compare and contrast coord_cartesian() vs xlim() or ylim() when zooming in on a histogram. What happens if you leave binwidth unset? What happens if you try and zoom so only half a bar shows?
```{r}

#coor_cartesian
ggplot(data = diamonds) + 
  geom_histogram(mapping = aes(x = price), binwidth = 100) +
  coord_cartesian(xlim = c(0, 10000))

#xlim()
ggplot(data = diamonds) +
  geom_histogram(mapping = aes(x = price), binwidth = 100) + xlim(0, 10000)

#coor_cartesian and xlim() generate the same shape when zooming in on a histogram. The difference lies on how they process the zoom-in command. coor_cartesian zoom in the data while xlim() removes the data that fall outside of the boundary.

#leave binwidth unset
ggplot(data = diamonds) + 
  geom_histogram(mapping = aes(x = price)) +
  coord_cartesian(xlim = c(0, 5000))
#When we leave binwidth unset, r studio uses bins = 30 as default setting.

#zoom so only half a bar shows
ggplot(data = diamonds) + 
  geom_histogram(mapping = aes(x = price), binwidth = 100) +
  coord_cartesian(xlim = c(0, 50))
#When the boundary contains only half a bar, rstudio does not show anything on the historgram.

```

7.4.1
1.What happens to missing values in a histogram? What happens to missing values in a bar chart? Why is there a difference?
```{r}

```


2.What does na.rm = TRUE do in mean() and sum()?
```{r}

```

7.5.1.1
1.Use what you’ve learned to improve the visualisation of the departure times of cancelled vs. non-cancelled flights.
```{r}

```


2.What variable in the diamonds dataset is most important for predicting the price of a diamond? How is that variable correlated with cut? Why does the combination of those two relationships lead to lower quality diamonds being more expensive?
```{r}

```


3.Install the ggstance package, and create a horizontal boxplot. How does this compare to using coord_flip()?
```{r}

```


4.One problem with boxplots is that they were developed in an era of much smaller datasets and tend to display a prohibitively large number of “outlying values”. One approach to remedy this problem is the letter value plot. Install the lvplot package, and try using geom_lv() to display the distribution of price vs cut. What do you learn? How do you interpret the plots?
```{r}

```


5.Compare and contrast geom_violin() with a facetted geom_histogram(), or a coloured geom_freqpoly(). What are the pros and cons of each method?
```{r}

```


6.If you have a small dataset, it’s sometimes useful to use geom_jitter() to see the relationship between a continuous and categorical variable. The ggbeeswarm package provides a number of methods similar to geom_jitter(). List them and briefly describe what each one does.
```{r}

```

Q3
Redo HW1 Q2 using tidyverse.
```{r}

```



