---
title: "hw2"
output: html_document
author: Lumi Huang
---
Q1
Read Chapter 7 (Exploratory Data Analysis) of R for Data Science and do exercises 7.3.4, 7.4.1, 7.5.1.1, 7.5.2.1, and 7.5.3.1

7.3.4
1.Explore the distribution of each of the x, y, and z variables in diamonds. What do you learn? Think about a diamond and how you might decide which dimension is the length, width, and depth.

```{r}


```

2.Explore the distribution of price. Do you discover anything unusual or surprising? (Hint: Carefully think about the binwidth and make sure you try a wide range of values.)
```{r}

```

3.How many diamonds are 0.99 carat? How many are 1 carat? What do you think is the cause of the difference?
```{r}

```


4.Compare and contrast coord_cartesian() vs xlim() or ylim() when zooming in on a histogram. What happens if you leave binwidth unset? What happens if you try and zoom so only half a bar shows?
```{r}

```

7.4.1
1.What happens to missing values in a histogram? What happens to missing values in a bar chart? Why is there a difference?
```{r}

```


2.What does na.rm = TRUE do in mean() and sum()?
```{r}

```

7.5.1.1
1.Use what you’ve learned to improve the visualisation of the departure times of cancelled vs. non-cancelled flights.
```{r}

```


2.What variable in the diamonds dataset is most important for predicting the price of a diamond? How is that variable correlated with cut? Why does the combination of those two relationships lead to lower quality diamonds being more expensive?
```{r}

```


3.Install the ggstance package, and create a horizontal boxplot. How does this compare to using coord_flip()?
```{r}

```


4.One problem with boxplots is that they were developed in an era of much smaller datasets and tend to display a prohibitively large number of “outlying values”. One approach to remedy this problem is the letter value plot. Install the lvplot package, and try using geom_lv() to display the distribution of price vs cut. What do you learn? How do you interpret the plots?
```{r}

```


5.Compare and contrast geom_violin() with a facetted geom_histogram(), or a coloured geom_freqpoly(). What are the pros and cons of each method?
```{r}

```


6.If you have a small dataset, it’s sometimes useful to use geom_jitter() to see the relationship between a continuous and categorical variable. The ggbeeswarm package provides a number of methods similar to geom_jitter(). List them and briefly describe what each one does.
```{r}

```

Q3
Redo HW1 Q2 using tidyverse.
```{r}

```



