---
title: "Biostat M280 Homework 3"
subtitle: Due Mar 2 @ 11:59PM
output: html_document
author: Lumi Huang
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Q1 LA City Employee Payroll

Link to my shiny app: <https://lumi-huang.shinyapps.io/hw3-shiny/>

The `/home/m280-data/la_payroll/LA_City_Employee_Payroll.csv` file on teaching server contains payroll information of LA City employees in years 2013-2017. It was downloaded from [LA City Controller's Office](https://controllerdata.lacity.org/Payroll/City-Employee-Payroll/pazn-qyym). Make a Shiny app to facilitate exploratory data analysis. 

1. For efficiency of the Shiny app, you should first pre-process, pare down, tidy, and save the data, e.g., as a compressed RDS file, to be used in the app.

0. **Total payroll by LA City**. Visualize the total LA City payroll of each year, with breakdown into base pay, overtime pay, and other pay.

0. **Who earned most?** Visualize the payroll information (total payment with breakdown into base pay, overtime pay, and other pay, Department, Job Title) of the top $n$ highest paid LA City employees in a specific year. User specifies $n$ (default 10) and year (default 2017).

0. **Which departments earn most?** Visualize the mean or median payroll, with breakdown into base pay, overtime pay, and other pay, of top $n$ earning departments. User specifies $n$ (default 5), year (default 2017), and method (mean or median, default median).

0. **Which departments cost most?** Visualize the total payroll, with breakdown into base pay, overtime pay, and other pay, of top $n$ expensive departments. User specifies $n$ (default 5) and year (default 2017).

0. Visualize any other information you are interested in.

0. Publish your Shiny app to <https://www.shinyapps.io> and share the link.



## Q2 LA City Parking War

The SQLite database `/home/m280-data/la_parking/LA_Parking_Citations.sqlite` on teaching server contains information about parking tickets in LA City. It was downloaded from [LA Open Data Portal](https://data.lacity.org/A-Well-Run-City/Parking-Citations/wjz9-h9np). Connect to the database and answer following questions using plots and summary statistics. In this exercise, you are **not** allowed to load whole data into memory. Use the _transform in database, plot in R_ strategy.
```{r}
library("DBI")
library("RSQLite")
library("tidyverse")
library("dbplyr")
setwd(".")
db <- dbConnect(RSQLite::SQLite(), dbname = 
                  "/home/m280-data/la_parking/LA_Parking_Citations_Extra.sqlite"
                )

dbListTables(db)
knitr::opts_chunk$set(connection = "db")

latix_sql <- dplyr::tbl(db, "latix")
str(latix_sql)
latix_sql %>% print(width = Inf)

```



1. How many tickets are in this data set?  
```{r}
latix_sql %>%
  count()
```


```{sql connection="db"}
select count(*) from latix;

```
Which time period do these tickets span?
Time span from 2010/04/28 to 2017/12/30.
```{r}
time_period <- latix_sql %>%
  select(Issue_Year, Issue_Month, Issue_Day) %>%
  group_by(Issue_Year, Issue_Month, Issue_Day) %>%
  distinct() %>%
  arrange(Issue_Year, Issue_Month, Issue_Day)

time_period1 <- head(time_period)
time_period1

time_period2 <- latix_sql %>%
  select(Issue_Year, Issue_Month, Issue_Day) %>%
  group_by(Issue_Year, Issue_Month, Issue_Day) %>%
  distinct() %>%
  arrange(desc(Issue_Year), desc(Issue_Month), desc(Issue_Day))

time_period3 <- head(time_period2)
time_period3


```


```{sql connection="db"}
select min(Issue_Year), max(Issue_Year) from latix;
```
```{sql connection="db"}
select min(Issue_Month) from latix
where Issue_Year = '2010';

```
```{sql connection="db"}
select min(Issue_Day) from latix
where Issue_Year = '2010' and Issue_Month = '4';
```
```{sql connection="db"}
select max(Issue_Month) from latix
where Issue_Year = '2017';

```

```{sql connection="db"}
select max(Issue_Day) from latix
where Issue_Year = '2017' and Issue_Month = '12';
```
tickets span from 2010/04/27 to 2017/12/30

Which years have most data?
```{r}
latix_sql %>%
  select(Issue_Year) %>%
  group_by(Issue_Year) %>%
  count() %>%
  arrange(desc(n))
```


```{sql connection="db"}
select Issue_Year, count(*) from latix
group by Issue_Year
order by count(*) desc;
```
2015 has most data


0. When (which hour, weekday, month day, and month) are you most likely to get a ticket and when are you least likely to get a ticket?
```{r}
#Most likely
#month
latix_sql %>%
  select(Issue_Month) %>%
  group_by(Issue_Month) %>%
  count() %>%
  arrange(desc(n))
#day
latix_sql %>%
  select(Issue_Day) %>%
  group_by(Issue_Day) %>%
  count() %>%
  arrange(desc(n))
#Weekday
latix_sql %>%
  select(Issue_Wday) %>%
  group_by(Issue_Wday) %>%
  count() %>%
  arrange(desc(n))

#Hour
latix_sql %>%
  select(Issue_Hour) %>%
  group_by(Issue_Hour) %>%
  count() %>%
  arrange(desc(n))

#Least Likely
#Month
latix_sql %>%
  select(Issue_Month) %>%
  group_by(Issue_Month) %>%
  count() %>%
  arrange(n)
#day
latix_sql %>%
  select(Issue_Day) %>%
  group_by(Issue_Day) %>%
  count() %>%
  arrange(n)
#Weekday
latix_sql %>%
  select(Issue_Wday) %>%
  group_by(Issue_Wday) %>%
  count() %>%
  arrange(n)

#Hour
latix_sql %>%
  select(Issue_Hour) %>%
  group_by(Issue_Hour) %>%
  count() %>%
  arrange(n)
  
```
Most Likely:
Month: 3
Day: 22
Weekday: Wed
Hour: 12pm

Least Likely:
Month: 11
Day: 31
Weekday: Sun
Hour: 5



0. Which car makes received most citations?

```{r}
latix_sql %>%
  select(Make) %>%
  group_by(Make) %>%
  count() %>%
  arrange(desc(n))
```


```{sql connection="db"}
select make, count(*) from latix
group by make
order by count(*) desc
limit 5;
```

Toyota

0. How many different colors of cars were ticketed? Which color attracted most tickets?
```{r}
latix_sql %>% 
  select(Color) %>%
  distinct() %>%
  count()

latix_sql %>%
  select(Color) %>%
  group_by(Color) %>%
  count() %>%
  arrange(desc(n))

```



There are 66 different colors of cars were ticketed, and color Black attracted most tickets.

0. What are the most common ticket types?
```{r}
latix_sql %>%
  select(Violation_Description) %>%
  group_by(Violation_Description) %>%
  count() %>%
  arrange(desc(n))
```
NO PARK/STREET CLEAN is the most common ticket types.

0. How much money was collected on parking tickets in 2015 and 2016?
```{r}
latix_sql %>%
  select(Issue_Year, Fine_amount) %>%
  filter(Issue_Year == "2015" | Issue_Year == "2016") %>%
  group_by(Issue_Year) %>%
  summarise(Total_Fine = sum(Fine_amount, na.rm = TRUE)) %>%
  arrange(desc(Total_Fine))
```

In 2015, 151006794 usd was collected. In 2016, 123236136 usd was collected.

0. Visualize any other information you are interested in.
Which location collect the most money in 2015
```{r}
latix_sql %>%
  select(Issue_Year, Location, Fine_amount) %>%
  filter(Issue_Year == "2015") %>%
  group_by(Location) %>%
  summarise(Total_Fine = sum(Fine_amount, na.rm = TRUE)) %>%
  arrange(desc(Total_Fine))
```


11600 SAN VICENTE BL collected the most money in 2015
