---
title: "Biostat M280 Homework 3"
subtitle: Due Mar 2 @ 11:59PM
output: html_document
author: Lumi Huang
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Q1 LA City Employee Payroll

The `/home/m280-data/la_payroll/LA_City_Employee_Payroll.csv` file on teaching server contains payroll information of LA City employees in years 2013-2017. It was downloaded from [LA City Controller's Office](https://controllerdata.lacity.org/Payroll/City-Employee-Payroll/pazn-qyym). Make a Shiny app to facilitate exploratory data analysis. 

1. For efficiency of the Shiny app, you should first pre-process, pare down, tidy, and save the data, e.g., as a compressed RDS file, to be used in the app.

0. **Total payroll by LA City**. Visualize the total LA City payroll of each year, with breakdown into base pay, overtime pay, and other pay.

0. **Who earned most?** Visualize the payroll information (total payment with breakdown into base pay, overtime pay, and other pay, Department, Job Title) of the top $n$ highest paid LA City employees in a specific year. User specifies $n$ (default 10) and year (default 2017).

0. **Which departments earn most?** Visualize the mean or median payroll, with breakdown into base pay, overtime pay, and other pay, of top $n$ earning departments. User specifies $n$ (default 5), year (default 2017), and method (mean or median, default median).

0. **Which departments cost most?** Visualize the total payroll, with breakdown into base pay, overtime pay, and other pay, of top $n$ expensive departments. User specifies $n$ (default 5) and year (default 2017).

0. Visualize any other information you are interested in.

0. Publish your Shiny app to <https://www.shinyapps.io> and share the link.

## Q2 LA City Parking War

The SQLite database `/home/m280-data/la_parking/LA_Parking_Citations.sqlite` on teaching server contains information about parking tickets in LA City. It was downloaded from [LA Open Data Portal](https://data.lacity.org/A-Well-Run-City/Parking-Citations/wjz9-h9np). Connect to the database and answer following questions using plots and summary statistics. In this exercise, you are **not** allowed to load whole data into memory. Use the _transform in database, plot in R_ strategy.
```{r}
library("DBI")
library("RSQLite")
library("tidyverse")
setwd(".")
db <- dbConnect(RSQLite::SQLite(), dbname = 
                  "/home/m280-data/la_parking/LA_Parking_Citations_Extra.sqlite")
dbListTables(db)
knitr::opts_chunk$set(connection = "db")

latix_sql <- dplyr::tbl(db, "latix")
str(latix_sql)
latix_sql %>% print(width = Inf)

```

```{sql connection="db"}
select * from latix
limit 5

```

1. How many tickets are in this data set?  


```{sql connection="db"}
select count(*) from latix;

```
Which time period do these tickets span?
```{sql connection="db"}
select min(Issue_Year), max(Issue_Year) from latix;
```
```{sql connection="db"}
select min(Issue_Month) from latix
where Issue_Year = '2010';

```
```{sql connection="db"}
select min(Issue_Day) from latix
where Issue_Year = '2010' and Issue_Month = '4';
```
```{sql connection="db"}
select max(Issue_Month) from latix
where Issue_Year = '2017';

```

```{sql connection="db"}
select max(Issue_Day) from latix
where Issue_Year = '2017' and Issue_Month = '12';
```
tickets span from 2010/04/27 to 2017/12/30

Which years have most data?
```{sql connection="db"}
select Issue_Year, count(*) from latix
group by Issue_Year
order by count(*) desc;
```
2015 has most data


0. When (which hour, weekday, month day, and month) are you most likely to get a ticket and when are you least likely to get a ticket?
```{sql connection="db"}
select Issue_Month, Issue_Day, Issue_Wday, Issue_Hour, count(*) from latix
group by Issue_Month, Issue_Day, Issue_Wday, Issue_Hour
order by count(*) desc
limit 5;
```
```{sql connection="db"}
select Issue_Month, Issue_Day, Issue_Wday, Issue_Hour, count(*) from latix
group by Issue_Month, Issue_Day, Issue_Wday, Issue_Hour
order by count(*)
limit 5;
```
Combined: Most likely on January 2nd (Sat) at 8am / Least likely on January 1st (Thu) 13pm (ties).

```{sql connection="db"}
select Issue_Month, count(*) from latix
group by Issue_Month
order by count(*) desc
limit 2;
```
```{sql connection="db"}
select Issue_Month, count(*) from latix
group by Issue_Month
order by count(*)
limit 2;
```
Most likely month: March, Least likely month: November

```{sql connection="db"}
select Issue_Day, count(*) from latix
group by Issue_Day
order by count(*) desc
limit 2;
```

```{sql connection="db"}
select Issue_Day, count(*) from latix
group by Issue_Day
order by count(*)
limit 2;
```
Most likely day: 22, Least likely day: 31

```{sql connection="db"}
select Issue_Wday, count(*) from latix
group by Issue_Wday
order by count(*) desc
limit 2;
```
```{sql connection="db"}
select Issue_Wday, count(*) from latix
group by Issue_Wday
order by count(*)
limit 2;
```
Most likely weekday: Wed, Least likely weekday: Sunday

```{sql connection="db"}
select Issue_Hour, count(*) from latix
group by Issue_Hour
order by count(*) desc
limit 2;
```
```{sql connection="db"}
select Issue_Hour, count(*) from latix
group by Issue_Hour
order by count(*)
limit 2;
```
Most likely hour: 12pm, Least likely hour: 5am

0. Which car makes received most citations?

```{sql connection="db"}
select make, count(*) from latix
group by make
order by count(*) desc
limit 5;
```

Toyota

0. How many different colors of cars were ticketed? Which color attracted most tickets?
```{sql connection="db"}
select count(distinct(color)) from latix;
```

```{sql connection="db"}
select color, count(*) from latix;
group by color
order by count(*) desc;
```
There are 65 different colors of cars were ticketed, and color SL attracted most tickets.

0. What are the most common ticket types?
```{sql connection="db"}

```


0. How much money was collected on parking tickets in 2015 and 2016?

0. Visualize any other information you are interested in.